---
title: "Troubleshooting Guide - NHANES BMI Body Fat Analysis"
author: "NHANES BMI Body Fat Analysis Team"
output:
  learnr::tutorial:
    progressive: true
    allow_skip: true
    css: css/styles.css
runtime: shiny
---

```{r setup, include=FALSE}
library(learnr)
library(yaml)
library(dplyr)

# Set knitr options for learnr
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

## Troubleshooting Guide 🛠️

Welcome to the comprehensive troubleshooting guide for the NHANES BMI Body Fat Analysis! This interactive guide will help you diagnose and resolve common issues.

### Quick Health Check

Let's start by checking your pipeline health:

```{r health-check, exercise=TRUE}
# Run the pipeline health check
tryCatch({
  source("../R/error_handling.R")
  display_pipeline_health()
}, error = function(e) {
  cat("❌ Could not run health check. Please check your R environment.\n")
  cat("Error:", e$message, "\n")
})
```

## Common Issues & Solutions

### 1. Installation Problems

#### ❌ "Package 'X' not found" errors

**Symptoms:**
- Error messages about missing packages during installation or runtime
- `library(package_name)` fails

**Solutions:**

1. **Install missing packages:**
```r
install.packages(c('dplyr', 'ggplot2', 'survey', 'foreign', 'yaml', 'future', 'furrr'))
```

2. **Update R and RStudio:**
```r
# In R console:
install.packages("installr")
installr::updateR()
```

3. **Check system dependencies:**
```bash
# On macOS with Homebrew:
brew install gfortran
```

4. **Use renv for reproducible environments:**
```r
# If renv is available:
renv::restore()
```

#### Quiz: Package Installation

```{r quiz-packages}
quiz(caption = "Package Installation Troubleshooting",
  question("What should you do first if you get a 'package not found' error?",
    answer("Reinstall R completely", message = "Try installing the specific package first."),
    answer("Install the missing package", correct = TRUE),
    answer("Delete all packages", message = "Install the missing package, don't delete others."),
    answer("Restart your computer", message = "Package installation should resolve the issue.")
  ),
  question("Which command installs multiple packages at once?",
    answer("install.packages(package1, package2)", message = "Use a character vector: c('package1', 'package2')."),
    answer("install.packages(c('package1', 'package2'))", correct = TRUE),
    answer("install.packages('package1', 'package2')", message = "Use a character vector: c('package1', 'package2')."),
    answer("install.packages(package1 + package2)", message = "Use a character vector: c('package1', 'package2').")
  )
)
```

### 2. Data Download Issues

#### ❌ "File not found" or "Data download failed" errors

**Symptoms:**
- `make fetch` fails
- Missing files in `data/raw/` directory
- Network timeout errors

**Solutions:**

1. **Check internet connection:**
```bash
# Test connection
ping -c 3 google.com
```

2. **Verify disk space:**
```bash
# Check available space
df -h
```

3. **Manual download (alternative):**
```bash
# Download files manually from CDC website:
# https://wwwn.cdc.gov/Nchs/Nhanes/2017-2018/DEMO_J.XPT
# Place in data/raw/ directory
```

4. **Retry with verbose output:**
```r
# Run with detailed logging
R -e "source('scripts/fetch_nhanes.R'); cat('Download completed\n')"
```

### 3. Configuration Problems

#### ❌ "Configuration file not found" or "Invalid YAML" errors

**Symptoms:**
- Missing `config/config.yml` file
- YAML parsing errors
- Missing required configuration sections

**Solutions:**

1. **Use the Configuration Wizard:**
```r
# Launch the interactive configuration tool
R -e "shiny::runApp('app.R')"
```

2. **Create config directory and file:**
```bash
mkdir -p config
# The wizard will create the config.yml file
```

3. **Validate YAML syntax:**
```yaml
# Check your config.yml file for proper indentation and structure
data:
  raw_dir: "data/raw"
  derived_dir: "data/derived"
# Use spaces, not tabs!
```

#### Interactive Exercise: Fix Configuration

Let's check and fix common configuration issues:

```{r config-check, exercise=TRUE}
# Check if config file exists
config_file <- "../config/config.yml"

if (file.exists(config_file)) {
  tryCatch({
    config <- yaml::read_yaml(config_file)
    cat("✅ Configuration file exists and is valid!\n")

    # Check required sections
    required_sections <- c("data", "outputs", "nhanes", "analysis", "logging")
    missing_sections <- required_sections[!required_sections %in% names(config)]

    if (length(missing_sections) == 0) {
      cat("✅ All required configuration sections present!\n")
    } else {
      cat("❌ Missing sections:", paste(missing_sections, collapse = ", "), "\n")
      cat("💡 Use the Configuration Wizard to fix this.\n")
    }
  }, error = function(e) {
    cat("❌ Configuration file has syntax errors.\n")
    cat("Error:", e$message, "\n")
    cat("💡 Use the Configuration Wizard to create a valid config file.\n")
  })
} else {
  cat("❌ Configuration file not found.\n")
  cat("💡 Run: R -e 'shiny::runApp(\"app.R\")' to create it.\n")
}
```

### 4. Directory Structure Issues

#### ❌ "Directory not found" or "Permission denied" errors

**Symptoms:**
- Missing directories in error messages
- Cannot write to output directories
- File creation fails

**Solutions:**

1. **Create required directories:**
```bash
# Create all required directories at once
mkdir -p data/raw data/derived outputs/tables outputs/figures outputs/logs outputs/report config cache tutorials
```

2. **Check permissions:**
```bash
# Check current directory permissions
ls -la

# Fix permissions if needed
chmod 755 data outputs config
```

3. **Use absolute paths in configuration:**
```yaml
# In config/config.yml, use absolute paths if relative paths fail
data:
  raw_dir: "/full/path/to/project/data/raw"
```

### 5. Memory and Performance Issues

#### ❌ "Out of memory" or "Process killed" errors

**Symptoms:**
- R crashes during analysis
- High memory usage
- Slow performance

**Solutions:**

1. **Use parallel processing:**
```bash
# Use the parallel pipeline for better performance
make parallel-pipeline
```

2. **Increase R memory limit:**
```r
# In R console:
memory.limit(size = 8192)  # Increase to 8GB
```

3. **Monitor system resources:**
```bash
# Check memory usage
top -o mem

# Check disk space
df -h
```

4. **Use caching:**
```bash
# Enable caching to avoid re-computation
make parallel-pipeline  # Automatically uses caching
```

### 6. Statistical Analysis Issues

#### ❌ "Convergence failed" or "Invalid results" errors

**Symptoms:**
- Survey analysis fails to converge
- Strange statistical results
- Warning messages about sample sizes

**Solutions:**

1. **Check sample sizes:**
```r
# Check sample size in your data
summary(your_data$sample_size_variable)
```

2. **Review survey design:**
```r
# Check survey weights distribution
summary(your_data$WTMEC2YR)
```

3. **Validate data quality:**
```r
# Check for missing values
colSums(is.na(your_data))
```

## Advanced Troubleshooting

### Debugging Tools

#### 1. Enable Debug Logging

```r
# In config/config.yml, set logging level to DEBUG
logging:
  level: "DEBUG"
  file: "analysis_log.txt"
```

#### 2. Run Step-by-Step

```bash
# Run individual pipeline steps for debugging
make fetch        # Download data
make cleandata    # Process data
make analysis     # Run analysis
make viz         # Generate plots
```

#### 3. Check Error Logs

```r
# View detailed error logs
cat outputs/logs/analysis_log.txt
cat outputs/logs/error_details.txt
```

### Interactive Debugging Exercise

Let's practice debugging a common issue:

```{r debug-exercise, exercise=TRUE}
# Simulate a common error scenario
# Uncomment and run this code to see error handling in action

# Example: Missing data file
# tryCatch({
#   data <- read.csv("nonexistent_file.csv")
# }, error = function(e) {
#   source("../R/error_handling.R")
#   fake_error <- NhanesError(
#     message = "File not found: nonexistent_file.csv",
#     code = "FILE_NOT_FOUND",
#     details = list(file_path = "nonexistent_file.csv")
#   )
#   display_user_friendly_error(fake_error)
# })

cat("💡 Tip: The enhanced error handling shows actionable suggestions for each error type.\n")
```

## Getting Help

### Self-Service Resources

1. **Interactive Tutorial**: `tutorials/getting_started.Rmd`
2. **Configuration Wizard**: `R -e 'shiny::runApp("app.R")'`
3. **Pipeline Health Check**: `R -e 'source("R/error_handling.R"); display_pipeline_health()'`
4. **Error Logs**: `outputs/logs/`

### Community Support

1. **GitHub Issues**: https://github.com/altalanta/nhanes-bmi-bodyfat/issues
2. **Documentation**: outputs/report/report.html
3. **Email Support**: analysis@nhanes-bmi.org

### Professional Support

For research collaborations or advanced customization:
- **Consulting Services**: Available for custom analysis and methodology
- **Training Workshops**: Hands-on training for research teams
- **Code Review**: Professional review of analysis pipelines

## Prevention Best Practices

### 1. Regular Health Checks

```r
# Run this before starting new analyses
source("R/error_handling.R")
display_pipeline_health()
```

### 2. Version Control

```bash
# Track your configuration and results
git add config/config.yml outputs/
git commit -m "Updated configuration for new analysis"
```

### 3. Documentation

```r
# Document your analysis parameters
# Create a README in your outputs directory
writeLines(c(
  "Analysis Date: ", Sys.Date(),
  "Configuration: See config/config.yml",
  "Results: See outputs/tables/ and outputs/figures/"
), "outputs/README.txt")
```

## Final Quiz: Troubleshooting Skills

```{r final-troubleshooting-quiz}
quiz(caption = "Troubleshooting Assessment",
  question("What should you do first when encountering an error?",
    answer("Panic and restart everything", message = "Stay calm and check the error message details."),
    answer("Read the error message carefully", correct = TRUE),
    answer("Delete all files and start over", message = "Check the error message and use the suggestions provided."),
    answer("Ignore it and hope it goes away", message = "Address errors systematically using the provided suggestions.")
  ),
  question("Which tool helps diagnose setup issues before running analysis?",
    answer("Pipeline health check", correct = TRUE),
    answer("Random guessing", message = "Use the systematic health check function."),
    answer("Crystal ball", message = "Use the systematic health check function."),
    answer("Magic wand", message = "Use the systematic health check function.")
  ),
  question("What's the best way to get help for complex issues?",
    answer("GitHub Issues page", correct = TRUE),
    answer("Random internet search", message = "Use the official GitHub issues for structured support."),
    answer("Yelling at the computer", message = "Use structured support channels."),
    answer("Hiring a psychic", message = "Use structured support channels.")
  )
)
```

## Congratulations! 🎉

You've completed the troubleshooting guide! You now have the skills to:

- ✅ Diagnose common NHANES analysis issues
- ✅ Apply systematic troubleshooting approaches
- ✅ Use enhanced error messages effectively
- ✅ Prevent issues through best practices
- ✅ Seek help when needed

### Remember:
- **Stay calm** when errors occur - most have simple solutions!
- **Check the error suggestions** - they're tailored to each error type
- **Use the health check** before starting new analyses
- **Document your solutions** for future reference

Happy troubleshooting! 🔧✨



